Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-01-22 PAPER.tex
Added some responses to Danny questions + added some comments to Appendix B

Using this motivation, the numerator of Equation \eqref{eq:Deriv2} is precisely our expression for $P_{\rm out}$ (Equation \eqref{eq:sigloss}), and the denominator is $P_{\rm in}$ (Equation \eqref{eq:Pin}), where the function $g$ is our OQE power spectrum pipeline. \dcj{Here is the post-hoc justification for this signal loss method.  Except goes back to assumption of equation for g which just \emph{assumes} that there are no cross terms in the estimation of a power spectrum. So we're just kicking the can down the road.} \acl{Ok, in my defense, Equation \eqref{eq:LinearPspecSum} is a little better than an \emph{assumption}. Appendix \eqref{sec:sigloss_appendix} does in fact illustrate this with an example. The crucial question is whether we're allowed to motivate our method based on a relationship that is only true in expectation. Note that this expectation includes allowances for signal loss due to the cross correlation between signal and foregrounds, with an example of this being Appendix \ref{sec:sigloss_appendix}.}
\jp{I'm fine with leaving the text as is and calling this note closed with the addition of the appendix, but will wait for sign off from DCJ.}
\dcj{ Assuming that expansion to "leading order in $\eta$" means $\eta<<1$ then the approximation is that the 
covariance is nearly perfectly calculated from the data with only small effects due to finite samples. Thus 
correlations don't make it into the expression and foregrounds remain uncorrelated with eor or injected signals.  
I added words explaining this but it still looks strange for us to be saying that loss is linear in one section and 
then explaining in the next section exactly how cross terms were what messed us up in the first go round.}

@@

The treatment we outlined in the previous section implicitly assumed that signal loss corrections could be treated in expectation. In practice, however, the signal loss itself is a random variable --- some realizations of the cosmological signal may be more correlated with foreground modes than others, leading to more signal loss. This leads to two issues. The first is that Equation \eqref{eq:LinearPspecSum} may not hold. In particular, a third term that is a quadratic combination of $\f$ and $\s$ (e.g., $\f^\dagger \C^{-1} \s$) may appear. However, we can circumvent this issue by decomposing $\f$ into a sum of two vectors, one that is proportional to $\s$ and one that is orthogonal to $\s$. The former can be absorbed into the EoR term (part of the signal loss estimate), while the latter can be absorbed into the foreground term. In other words, suppose that we have
\begin{equation}
\phat_x \approx \ell_{\rm fg} |\mathbf{f}|^2 + \ell_{\rm eor} |\mathbf{e}|^2 + \ell_{\rm fe} \mathbf{f} \cdot \mathbf{e}.
\end{equation}
in place of Equation \eqref{eq:LinearPspecSum}. Now, split $\mathbf{f}$ into $\mathbf{ f}_\parallel +\mathbf{ f}_\perp$, which are the components of the foregrounds that couple to the EoR versus those that do not. Further parameterize $\mathbf{ f}_\parallel \equiv \rho \mathbf{e}$, where $\rho$ is some correlation coefficient that quantifies the degree of correlation between the foregrounds and the EoR. This then gives
\begin{equation}
\label{eq:fperpfpara}
\phat_x \approx \ell_{\rm fg} |\mathbf{f}|^2 +( \ell_{\rm eor} + \rho  \ell_{\rm fe} ) |\mathbf{e}|^2 + \ell_{\rm fe} \mathbf{f}\cdot \mathbf{f}_\perp.
\end{equation}
Since $|\mathbf{e}|^2$ would be a lossless estimate of the true EoR power spectrum, this expression recovers the linear relation between the true EoR power spectrum and the lossy estimate.
\dcj{Wait what? I think you are saying essentially that some x can remain in Pout which is what we are seeing. Or is this some extra bit of method which has been added to the analysis...} \acl{I'm not sure that's what I'm saying. (And by ``I'm not sure" I don't mean it in the colloquial sense of ``you're not right and I'm just trying to be polite about it", but rather the literal sense of the phrase ``I'm not sure"). What I was trying to say was that if I wrote Equation \eqref{eq:LinearPspecSum} without the ensemble average, I could say
\begin{equation}
\phat_x \approx \ell_{\rm fg} f^2 + \ell_{\rm eor} e^2 + \ell_{\rm fe} fe.
\end{equation}
(Note that even this is an approximation. In principle there is an infinite series containing all possible powers of 
$e$ and $f$). Now, we split $f$ into $ f_\parallel + f_\perp$, which are the components of the foregrounds that 
couple to the EoR versus those that do not. Further parameterize $f_\parallel \equiv \rho e$, where $\rho$ is 
some correlation coefficient that quantifies the degree of correlation between the foregrounds and the EoR. 
Under this parameterization, one can write the $fe$ term in terms of $e^2$ (which gets absorbed into the other 
EoR term, plus residual foregrounds that we don't care about).}\acl{Just promoted this last part to something in 
the main text.}\dcj{I will admit to being a little lost here.  Doesn't changing the basis just convert the unknown f 
dot s term into other unknown terms. Isn't the whole point of Eq 24 supposed to be that we can subtract off the 
$g(x)$ to estimate $\hat{\ell}_{\rm eor}$?  How do we subtract off $f\dot f_\perp$? Don't we want to know $
\ell_{\rm eor}$ not $(\ell_{\rm eor} + \rho\ell_{\rm fe})$? }

@@

\section{Simplified Loss Correction}
\label{sec:Pr_appendix}
One of the main difficulties in defining a signal loss procedure is in the interpretation of the injection procedure `output'. The method described in Section \ref{sec:Sigloss} relies on several assumptions which required justification including the assumption that the number of samples is large enough to treat sample variance as a first order purturbation on the ideal covariance matrix instead of the dominant effect we know it is. However it is possible to ask a slightly different question and with less math still get roughly the same answer.

In the main text we made an estimate of the signal loss effecting only the injected EoR by supposing that the lossy injected signal was the difference between the injected and not injected cases: $P_{out} = \widehat{\textbf{p}}_{r} - \widehat{\textbf{p}}_{x}$. Suppose that we rephrase the question we are asking to: ``At what point would an injected signal be detectable at significant levels?". Instead of disentangling the part of the output power spectrum dominated by the injected signal we can simply use $\widehat{\textbf{p}}_{r}$, the weighted power spectrum of the data plus injected signal.

@@

Formally we can express our test as: ``What is the $P_{\rm in}$ such that the power spectrum of "data + 
injected eor" ($\widehat{p}_{e+x}$) is detectable above the power spectrum of the data itself ($\widehat{p}_x$. 
In other words, what is $e$ such that,
\begin{equation}
\widehat{\textbf{p}}_{e+x} > \widehat{\textbf{p}}_{x},
\end{equation}  
?''  Of course, both sides of this equation are actually a distribution of values associated with data variance and injected model variance. One way to interpret this test for the two distributions is to compute the probability that the two distributions are not (i.e. the null test) the same.

 \begin{equation}
\Prob(P_{e})  =  1 -  \int \Prob(\widehat{\textbf{p}}_x) \Prob(\widehat{\textbf{p}}_{e+x}|P_{e}) x
 \end{equation}
 where $\Prob(\widehat{\textbf{p}}_x)$ is the probability of obtaining the power spectrum $\widehat{\textbf{p}}_x$ found via bootstrapping (power spectrum of data only) and $\Prob(\widehat{\textbf{p}}_{e+x}|P_{e})$ is the probability of obtaining the summed power spectrum $\widehat{\textbf{p}}_{x+e}$ given a range of input power levels, as sampled in Figure \ref{fig:Pr_vs_Pin}.

>>>>

Using this motivation, the numerator of Equation \eqref{eq:Deriv2} is precisely our expression for $P_{\rm out}$ (Equation \eqref{eq:sigloss}), and the denominator is $P_{\rm in}$ (Equation \eqref{eq:Pin}), where the function $g$ is our OQE power spectrum pipeline. \dcj{Here is the post-hoc justification for this signal loss method.  Except goes back to assumption of equation for g which just \emph{assumes} that there are no cross terms in the estimation of a power spectrum. So we're just kicking the can down the road.} \acl{Ok, in my defense, Equation \eqref{eq:LinearPspecSum} is a little better than an \emph{assumption}. Appendix \eqref{sec:sigloss_appendix} does in fact illustrate this with an example. The crucial question is whether we're allowed to motivate our method based on a relationship that is only true in expectation. Note that this expectation includes allowances for signal loss due to the cross correlation between signal and foregrounds, with an example of this being Appendix \ref{sec:sigloss_appendix}.}
\jp{I'm fine with leaving the text as is and calling this note closed with the addition of the appendix, but will wait for sign off from DCJ.}
\dcj{ Assuming that expansion to "leading order in $\eta$" means $\eta<<1$ then the approximation is that the 
covariance is nearly perfectly calculated from the data with only small effects due to finite samples. Thus 
correlations don't make it into the expression and foregrounds remain uncorrelated with eor or injected signals.  
I added words explaining this but it still looks strange for us to be saying that loss is linear in one section and 
then explaining in the next section exactly how cross terms were what messed us up in the first go round.} 
\acl{Ok, I'm open to suggestions as to how to phrase this, because lots of people seem to be thinking that the appendix was meant to justify everything. It's not. It's meant as a motivational example. I just wanted something where I could follow through analytically to the end to show that it's ok to think about signal loss as a multiplicative factor. At least in expectation. And then how one deals with not correcting for the signal loss in expectation is described in Section \ref{sec:Practice}.}

@@

The treatment we outlined in the previous section implicitly assumed that signal loss corrections could be treated in expectation. In practice, however, the signal loss itself is a random variable --- some realizations of the cosmological signal may be more correlated with foreground modes than others, leading to more signal loss. This leads to two issues. The first is that Equation \eqref{eq:LinearPspecSum} may not hold. In particular, a third term that is a quadratic combination of $\f$ and $\s$ (e.g., $\f^\dagger \C^{-1} \s$) may appear. However, we can circumvent this issue by decomposing $\f$ into a sum of two vectors, one that is proportional to $\s$ and one that is orthogonal to $\s$. The former can be absorbed into the EoR term (part of the signal loss estimate), while the latter can be absorbed into the foreground term. In other words, suppose that we have
\begin{equation}
\phat_x \approx \ell_{\rm fg} |\mathbf{f}|^2 + \ell_{\rm eor} |\mathbf{e}|^2 + \ell_{\rm fe} \mathbf{f} \cdot \mathbf{e}.
\end{equation}
in place of Equation \eqref{eq:LinearPspecSum}. Now, split $\mathbf{f}$ into $\mathbf{ f}_\parallel +\mathbf{ f}_\perp$, which are the components of the foregrounds that couple to the EoR versus those that do not. Further parameterize $\mathbf{ f}_\parallel \equiv \rho \mathbf{e}$, where $\rho$ is some correlation coefficient that quantifies the degree of correlation between the foregrounds and the EoR. This then gives
\begin{equation}
\label{eq:fperpfpara}
\phat_x \approx \ell_{\rm fg} |\mathbf{f}|^2 +( \ell_{\rm eor} + \rho  \ell_{\rm fe} ) |\mathbf{e}|^2 + \ell_{\rm fe} \mathbf{f}\cdot \mathbf{f}_\perp.
\end{equation}
Since $|\mathbf{e}|^2$ would be a lossless estimate of the true EoR power spectrum, this expression recovers the linear relation between the true EoR power spectrum and the lossy estimate.
\dcj{Wait what? I think you are saying essentially that some x can remain in Pout which is what we are seeing. Or is this some extra bit of method which has been added to the analysis...} \acl{I'm not sure that's what I'm saying. (And by ``I'm not sure" I don't mean it in the colloquial sense of ``you're not right and I'm just trying to be polite about it", but rather the literal sense of the phrase ``I'm not sure"). What I was trying to say was that if I wrote Equation \eqref{eq:LinearPspecSum} without the ensemble average, I could say
\begin{equation}
\phat_x \approx \ell_{\rm fg} f^2 + \ell_{\rm eor} e^2 + \ell_{\rm fe} fe.
\end{equation}
(Note that even this is an approximation. In principle there is an infinite series containing all possible powers of 
$e$ and $f$). Now, we split $f$ into $ f_\parallel + f_\perp$, which are the components of the foregrounds that 
couple to the EoR versus those that do not. Further parameterize $f_\parallel \equiv \rho e$, where $\rho$ is 
some correlation coefficient that quantifies the degree of correlation between the foregrounds and the EoR. 
Under this parameterization, one can write the $fe$ term in terms of $e^2$ (which gets absorbed into the other 
EoR term, plus residual foregrounds that we don't care about).}\acl{Just promoted this last part to something in 
the main text.}\dcj{I will admit to being a little lost here.  Doesn't changing the basis just convert the unknown f 
dot s term into other unknown terms. Isn't the whole point of Eq 24 supposed to be that we can subtract off the 
$g(x)$ to estimate $\hat{\ell}_{\rm eor}$?  How do we subtract off $f\dot f_\perp$? Don't we want to know $
\ell_{\rm eor}$ not $(\ell_{\rm eor} + \rho\ell_{\rm fe})$? }
\acl{I think what we are trying to say here is that actually the thing we want \emph{is} $(\ell_{\rm eor} + \rho\ell_{\rm fe})$. Essentially, the potential coupling between the injected EoR and the foregrounds means that we lose the nice form of Equation \eqref{eq:LinearPspecSum}, which justified what is basically a multiplicative correction to signal loss. What we are suggesting here is that even when there is a coupling, the estimator can be parameterized as $\alpha  |\mathbf{e}|^2 + \beta$. So essentially we are defining an effective $
\ell_{\rm eor}$, which is given by $(\ell_{\rm eor} + \rho\ell_{\rm fe})$.}

@@

\section{Simplified Loss Correction}
\label{sec:Pr_appendix}
One of the main difficulties in defining a signal loss procedure is in the interpretation of the injection procedure `output'. The method described in Section \ref{sec:Sigloss} relies on several assumptions which required justification including the assumption that the number of samples is large enough to treat sample variance as a first order purturbation on the ideal covariance matrix instead of the dominant effect we know it is. \acl{I'm not sure the previous sentence was fair. Again, the appendix is a motivating example. Carina's numerical computations do not rely on the derivation.} However it is possible to ask a slightly different question and with less math still get roughly the same answer.

In the main text we made an estimate of the signal loss effecting only the injected EoR by supposing that the lossy injected signal was the difference between the injected and not injected cases: $P_{out} = \widehat{\textbf{p}}_{r} - \widehat{\textbf{p}}_{x}$. Suppose that we rephrase the question we are asking to: ``At what point would an injected signal be detectable at significant levels?". \acl{Can we justify better why this is the right question? In particular, why is the detection of an injected signal the right thing to look at? After all, the universe doesn't do any signal injections. I think to justify that signal injections are the right thing to consider in the first place, one requires some sort of correspondence between equations analogous to Equation \eqref{eq:Deriv1} and \eqref{eq:Deriv2}.} Instead of disentangling the part of the output power spectrum dominated by the injected signal we can simply use $\widehat{\textbf{p}}_{r}$, the weighted power spectrum of the data plus injected signal.

@@

Formally we can express our test as: ``What is the $P_{\rm in}$ such that the power spectrum of "data + 
injected eor" ($\widehat{p}_{e+x}$) is detectable above the power spectrum of the data itself ($\widehat{p}_x$. 
In other words, what is $e$ such that,
\begin{equation}
\widehat{\textbf{p}}_{e+x} > \widehat{\textbf{p}}_{x},
\end{equation}  
?''  Of course, both sides of this equation are actually a distribution of values associated with data variance and injected model variance. One way to interpret this test for the two distributions is to compute the probability that the two distributions are not (i.e. the null test) the same.  
 \acl{I'm confused a little by this equation. I would've thought that the right thing to do would be to first compute $x_0$, given implicitly by
 \begin{equation}
 \int_{-\infty}^{x_0}\Prob(\widehat{\textbf{p}}_x) dx = \Lambda,
 \end{equation}
 where $\Lambda$ is the appropriate number for $2\sigma$. That tells us what $x$ would be needed for us to thing that it was an outlier result without there being an extra $e$ in the data. Then we ask how often we get such an outlier result by computing
  \begin{equation}
 \int_{x_0}^\infty \Prob(\widehat{\textbf{p}}_{e+x}|P_{e}) dx.
\end{equation}
Is what you've written equivalent to this? Sorry if this is remedial...
 }
 \begin{equation}
\Prob(P_{e})  =  1 -  \int \Prob(\widehat{\textbf{p}}_x) \Prob(\widehat{\textbf{p}}_{e+x}|P_{e}) x
 \end{equation}
 where $\Prob(\widehat{\textbf{p}}_x)$ is the probability of obtaining the power spectrum $\widehat{\textbf{p}}_x$ found via bootstrapping (power spectrum of data only) and $\Prob(\widehat{\textbf{p}}_{e+x}|P_{e})$ is the probability of obtaining the summed power spectrum $\widehat{\textbf{p}}_{x+e}$ given a range of input power levels, as sampled in Figure \ref{fig:Pr_vs_Pin}.
